A formal approach but practical approach in train regulation

### 2025-06-05 paper/main.tex
oups, une image oubliée

\begin{document}
%
\title{A formal approach in train regulation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Guillaume Bonfante \and Martin Vassor  \and Lucas Villaume}
%
\authorrunning{G. Bonfante et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Université de Lorraine -- CNRS -- LORIA}
%
\maketitle
%
\begin{abstract}
	...

@@

This contribution finds its origin in a cyber-warefare we are organising for the IT students at the University of Lorraine, see~\cite{CHE}. In the game-play, we have both IT infrastructures and physical systems, with connections between them. On the physical side, we built a small train network that tries to be a good model of real systems.  Our system manages customer informations, train regulation and finally rail infrastructure. Attackers can hit any of the three parts: disfiguring the site, changing a platform or blocking a signal. Defenders will protect their system, typically via a strict network policy.

@@

From an IT point of view, our system is shaped as shown on the right. Switches are controlled by a raspberry (TURNOUT) running an OpenPLC program, signals are controlled by an other raspberry and both communicate via modbus. The SIGNAL automaton reads mouvement sensors outputs (not all links shown in the figure). Next to that, we have a third machine, the regulator: it is in charge of updating train routes (as required by the customer IT),  of the regulation of turnouts and finally it may lock some signal (turn it to "red"). All the network connections are done via a single switch, connections from REGULATOR to TURNOUT and SIGNAL are done via modbus while the customer IT and the regulator talks  are based on TCP/IP.  To activate train mouvements, we use the DCC protocol via an Arduino, but this part is strictly related to the fact we have a model train.

\begin{wrapfigure}[7]{r}{0.3\textwidth}
\hspace{-10mm}
 \begin{minipage}{0.3\textwidth}
        \centering
        \vspace{-24mm}
\includegraphics[height=50mm]{ZoomAffichage.png}
    \end{minipage}
\end{wrapfigure}

@@

\begin{wrapfigure}[8]{r}{0.28\textwidth}
 \begin{minipage}{0.28\textwidth}
        \centering
        \vspace{-1mm}
\includegraphics[height=50mm]{maquette.png}
    \end{minipage}
\end{wrapfigure}

@@

 Now, we can enter in more details about how the train regulation works. First, for trains, we have a simple policy.  The regulator will ask the train to start in some direction (Forward/Backward). Then, the train will continue until a signal stops it and then restart when the signal turns green. It will do that up to its arrival point. On the other side, the regulator will, depending on the position of the trains, change switches position, lock/unlock some signal, ask for a train departure. All these operations are done by what we call an order. Trains Orders have the shape "Start direction Until position". Regulator orders are: XXX

  We do not make hypotheses on the durations of travels, but once a train is started and not stopped at some point, it will reach in finite time the next block.

@@

\subsection{Mosélisation d'une liste d'odre}
La modélisation de notre liste d'ordre doit prendre en compte les besoins et contraintes évoqués plus haut.
Pour cela, nous avons choisi de la représenter comme cela : $O = \{\Gamma, R\}$, avec $\Gamma$ l'ensemble des trains et $R$ notre régulateur.
Chaque train peut être représenté par un identifiant unique, une position, une direction et son programme : $t = (id, pos, dir, P)$, avec $t \in \Gamma$.
Le régulateur, portant une grande partie de la logique du modèle, se représente de manière plus complexe : $R = (E,T,S,W,G,H,TL)$.
%##########
%Définit ce qu'est un event à ce moment ou plus haut ?
%##########
Nous ne détaillerons pas l'ensemble des éléments de $R$ ici, mais nous pouvons en donner un aperçu rapide : $E$ contient l'ensemble des events,
$T$ le token courant pour chaque ressource et $H$ la dernière position connue de chaque train par le régulateur.

>>>>

\begin{document}
%
\title{A formal approach but practical approach in train regulation}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Guillaume Bonfante \and Martin Vassor  \and Lucas Villaume}
%
\authorrunning{G. Bonfante et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Université de Lorraine -- CNRS -- INRIA -- LORIA}
%
\maketitle
%
\begin{abstract}
	...

@@

This contribution finds its origin in a cyber-warefare we are organising for the IT students at the University of Lorraine, see~\cite{CHE}. In the game-play, we have both IT infrastructures and physical systems, with connections between them. On the physical side, we built a small train network that tries to be a good model of real systems.  Our system manages customer informations, train regulation and finally the physical/infrastructure layer. Attackers can hit any of the three parts: disfiguring the site, changing a platform or blocking a signal. Defenders will protect their system, typically via a strict network policy.

@@

From an IT point of view, our system is shaped as shown on the right. Switches are controlled by a raspberry (TURNOUT) running an OpenPLC program, signals are controlled by an other raspberry and SIGNAL has access to turnout positions via modbus. Then, the SIGNAL PLC reads mouvement sensors outputs (not all links shown in the figure) and updates accordingly the status of signals. Next to that, we have a third machine, the regulator: it is in charge of updating train routes (as required by the customer IT),  of the regulation of turnouts and finally it may lock some signal (turn it to "red"). All the network connections are done via a single switch, connections from REGULATOR to TURNOUT and SIGNAL are done via modbus while the customer IT and the regulator talks  are based on TCP/IP.  To activate train mouvements, we use the DCC protocol via an Arduino, but this part is strictly related to the fact we have a model train.

\begin{wrapfigure}[5]{r}{0.3\textwidth}
\hspace{-10mm}
 \begin{minipage}{0.3\textwidth}
        \centering
        \vspace{-24mm}
\includegraphics[height=50mm]{ZoomAffichage.png}
    \end{minipage}
\end{wrapfigure}

@@

\begin{wrapfigure}[8]{r}{0.28\textwidth}
 \begin{minipage}{0.28\textwidth}
        \centering
        \vspace{-10mm}
\includegraphics[height=50mm]{maquette.png}
    \end{minipage}
\end{wrapfigure}

@@

 Now, we can enter in more details about how the train regulation works. First, we have sensors on the rails at the two ends of each block. Accordingly, we make the hypothesis we can catch train entrances within blocks. Second, we make the hypothesis that switches can be swapped from one position to the other without fault. Our method could cope such issues, but we leave that for future work.

  Third, for trains, we have a simple policy.  According to its plan, it may start in some direction (Forward/Backward) up to some final position. Then, the train will run according to the signals: when green, it (re-)starts, when red, it stops.  We do not make hypotheses on the durations of travels, but once a train is running, it will reach in finite time the next block. Trains run in parallel. One train may leave a block, while an other is still in the middle of its own and the third train is breaking. There is no communications between the trains.

Fourth,  the regulator is capable of reading train block entrance events, changing some switch position and finally locking (that is make sure a signal is red)/unlocking some signal. We consider that those latter operations can be carried out without delay. Finally, the regulator may append some new duty to a train. To sum up, to control the train mouvements, the regulator can only lock/unlock some signal. Naturally, a signal intended to block train $T_1$ may block $T_2$ if not set at the right time.

 All these operations are done by what we call  \emph{orders}. Roughly speaking, trains Orders have the shape "Start in direction Until position". Regulator orders are: XXX.

 To sum up, the regulation policy can be seen as an asynchronous distributed system. The N trains and the regulator are independent agents.

 **** Comparaison avec les travaux antérieurs ****

Due to the distributed nature of the regulation, building a safe policy can be very tricky. We want to show in this paper that tools from formal methods are of great help.  In this paper, we describe in  a first step the formalisation process. We go up to a TLA+ specification that will allow us checking two properties of route orchestration: safety and liveness. This is our first research question.

Our second research question is the related to the following need. Suppose we want to append a new duty once a train reached its destination. This can be formalised as a composition problem (how to compose/concatenate some routes with some others). We describe such a composition procedure and then, again, we check that the routes are both safe and deadlock free.

@@

\subsection{Modélisation d'une liste d'ordre}
La modélisation de notre liste d'ordre doit prendre en compte les besoins et contraintes évoqués plus haut.
Pour cela, nous avons choisi de la représenter comme cela : $O = \{\Gamma, R\}$, avec $\Gamma$ l'ensemble des trains et $R$ notre régulateur.
Chaque train peut être représenté par un identifiant unique, une position, une direction et son programme : $t = (id, pos, dir, P)$, avec $t \in \Gamma$.
Le régulateur, portant une grande partie de la logique du modèle, se représente de manière plus complexe : $R = (E,T,S,W,G,H,TL)$.
%##########
%Définit ce qu'est un event à ce moment ou plus haut ?
%##########
Nous ne détaillerons pas l'ensemble des éléments de $R$ ici, mais nous pouvons en donner un aperçu rapide : $E$ contient l'ensemble des events,
$T$ le token courant pour chaque ressource et $H$ la dernière position connue de chaque train par le régulateur.
